Using Refactoring Techniques to Reduce Duplication in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-04-24 sattose/paper.tex
Talking about contents

\section{Introduction}
Duplication in source code is often seen as one of the most harmful types of technical debt. In Martin Fowler's ``Refactoring'' book \cite{fowler2018refactoring}, he exclaims that \textit{``Number one in the stink parade is duplicated code. If you see the same code structure in more than one place, you can be sure that your program will be better if you find a way to unify them.''}. In this research, we take a look at the challenges and opportunities in automatically refactoring duplicated code, also known as ``code clones''. The main goal is to improve the maintainability of the refactored code.

Refactoring is used to improve quality related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler2018refactoring, wake2004refactoring}, that are integrated into most modern IDE's. % May I assume the reader understands the word IDE?
However, most of these methods still require manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this research we will look at to what extend code clones can be automatically refactored.

@@

\subsection{Clone instance contents}
Finally, we looked at the contents of individual clone instances: what kind of declarations and statements do they span. We selected the following categories to be relevant for refactoring:
\begin{enumerate}
  \item \textbf{Full Method/Class/Interface/Enumeration/Constructor:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans a method partially, optionally including its declaration.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods (so not fields or anything in between).
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Includes Fields/Constructor:} A clone that spans a combination of fields and other things, like methods.
  \item \textbf{Method/Class/Interface/Enumeration Declaration:} A clone that contains the declaration (usually the first line) of a class, method, interface or enumeration.
  \item \textbf{Other:} Anything that does not match with above stated categories.
\end{enumerate}

@@

\begin{table}[H]
  \begin{center}
  \caption{Clone instance contents} \label{table:contents}
  \bigskip
\begin{tabular}{|l|l|l|}
  \hline
  \textbf{Contents}                                                        & \textbf{Amount} & \textbf{Percentage} \\ \hline
  Partial Method                                                           & 19.264          & 56,07               \\ \hline
  Several Methods                                                          & 9.076           & 26,41               \\ \hline
  Includes Constructor                                                     & 1.528           & 4,45                \\ \hline
  Includes Field                                                           & 1.149           & 3,34                \\ \hline
  Partial Constructor                                                      & 1.098           & 3,20                \\ \hline
  Only Fields                                                              & 565             & 1,64                \\ \hline
  Full Method                                                              & 554             & 1,61                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Class\\ Declaration\end{tabular}     & 445             & 1,30                \\ \hline
  Other                                                                    & 369             & 1,07                \\ \hline
  Full Class                                                               & 192             & 0,56                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Enum\\ Constant\end{tabular}         & 52              & 0,15                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Enum\\ Declaration\end{tabular}      & 47              & 0,14                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Interface\\ Declaration\end{tabular} & 10              & 0,03                \\ \hline
  Full Interface                                                           & 6               & 0,02                \\ \hline
  Full Enumeration                                                         & 4               & 0,01                \\ \hline
  Full Constructor                                                         & 1               & 0,00                \\ \hline
\end{tabular}
\end{center}
\end{table}
%TODO: Finish text and add measurements.

>>>>

\section{Introduction}
Duplication in source code is often seen as one of the most harmful types of technical debt. In Martin Fowler's ``Refactoring'' book \cite{fowler1999refactoring}, he exclaims that \textit{``Number one in the stink parade is duplicated code. If you see the same code structure in more than one place, you can be sure that your program will be better if you find a way to unify them.''}. In this research, we take a look at the challenges and opportunities in automatically refactoring duplicated code, also known as ``code clones''. The main goal is to improve the maintainability of the refactored code.

Refactoring is used to improve quality related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler1999refactoring, wake2004refactoring}, that are integrated into most modern IDE's. However, most of these methods still require manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this research we will look at to what extend code clones can be automatically refactored.

@@

\subsection{Clone instance contents}
Finally, we looked at the contents of individual clone instances: what kind of declarations and statements do they span. We selected the following categories to be relevant for refactoring:
\begin{enumerate}
  \item \textbf{Full Method/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans a method partially, optionally including its declaration.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods (so not fields or anything in between).
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Includes Fields/Constructor:} A clone that spans a combination of fields and other things, like methods.
  \item \textbf{Method/Class/Interface/Enumeration Declaration:} A clone that contains the declaration (usually the first line) of a class, method, interface or enumeration.
  \item \textbf{Other:} Anything that does not match with above stated categories.
\end{enumerate}

@@

\begin{table}[H]
  \begin{center}
  \caption{Clone instance contents} \label{table:contents}
  \bigskip
\begin{tabular}{|l|l|l|}
  \hline
  \textbf{Contents}                                                        & \textbf{Amount} & \textbf{Percentage} \\ \hline
  Partial Method                                                           & 19.264          & 56,07               \\ \hline
  Several Methods                                                          & 9.076           & 26,41               \\ \hline
  Includes Constructor                                                     & 1.528           & 4,45                \\ \hline
  Includes Field                                                           & 1.149           & 3,34                \\ \hline
  Partial Constructor                                                      & 1.098           & 3,20                \\ \hline
  Only Fields                                                              & 565             & 1,64                \\ \hline
  Full Method                                                              & 554             & 1,61                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Class\\ Declaration\end{tabular}     & 445             & 1,30                \\ \hline
  Other                                                                    & 369             & 1,07                \\ \hline
  Full Class                                                               & 192             & 0,56                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Enum\\ Constant\end{tabular}         & 52              & 0,15                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Enum\\ Declaration\end{tabular}      & 47              & 0,14                \\ \hline
  \begin{tabular}[c]{@{}l@{}}Includes Interface\\ Declaration\end{tabular} & 10              & 0,03                \\ \hline
  Full Interface                                                           & 6               & 0,02                \\ \hline
  Full Enumeration                                                         & 4               & 0,01                \\ \hline
\end{tabular}
\end{center}
\end{table}

Unsurprisingly, most clones span a part of a method. Just 1.6\% of the methods are cloned fully. More than a quarter of the clones spans over several methods, which is surprising. Simply extracting methods won't work in a case where a clone spans over several methods. About 4.5\% of clones include a constructor. About 3.3\% of clones include a global variable defined in a class.

\section{Redefining type-2 and type-3 clones}
